---
title: "gdistance::transition()"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 600, fig.width = 8)
```

`r Sys.Date()`

Understand the transition matrix, as used in `pathdistGen()`.

```{r, include=FALSE}
library(gdistance)
```


# Transition with `max()`

1. Find the maximum value for each pair of cell i and surrounding cell j.
2. Take the average of these values.

## Example 1

```{r}
BNG <- CRS("+proj=utm +zone=20 +ellps=GRS80 +units=m +no_defs")

r <- raster(xmn = 0, xmx = 4, ymn = 0, ymx = 4, res = 1, crs = BNG)
r[] <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)

plot(r)
text(r)
```

::: panel-tabset
### Directions: 4
```{r}
r_max <- transition(r, max, directions = 4, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 8
```{r}
r_max <- transition(r, max, directions = 8, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 16
```{r}
r_max <- transition(r, max, directions = 16, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

:::

## Example 2

```{r}
r <- raster(xmn = 0, xmx = 9, ymn = 0, ymx = 9, res = 1, crs = BNG)
r[] <- c(rep(0, 40), 1, rep(0, 40))

plot(r)
text(r)
```

::: panel-tabset
### Directions: 4
```{r}
r_max <- transition(r, max, directions = 4, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 8
```{r}
r_max <- transition(r, max, directions = 8, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 16
```{r}
r_max <- transition(r, max, directions = 16, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

:::

## Example 3

```{r}
r <- raster(xmn = 0, xmx = 4, ymn = 0, ymx = 4, res = 1, crs = BNG)
r[] <- c(0, 2, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)

plot(r)
text(r)
```

::: panel-tabset
### Directions: 4
```{r}

r_max <- transition(r, max, directions = 4, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 8
```{r}
r_max <- transition(r, max, directions = 8, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 16
```{r}
r_max <- transition(r, max, directions = 16, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

:::

# Transition with `1 / max()`

## Example 1

```{r}
BNG <- CRS("+proj=utm +zone=20 +ellps=GRS80 +units=m +no_defs")

r <- raster(xmn = 0, xmx = 4, ymn = 0, ymx = 4, res = 1, crs = BNG)
r[] <- c(1, 1, 1, 1, 1, 2, rep(1, 10))

plot(r)
text(r)
```

::: panel-tabset
### Directions: 4
```{r}
r_max <- transition(r, function(x) 1/max(x), directions = 4, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 8
```{r}
r_max <- transition(r, function(x) 1/max(x), directions = 8, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

### Directions: 16
```{r}
r_max <- transition(r, function(x) 1/max(x), directions = 16, symm = FALSE)

plot(raster(r_max), cex = 0.8)
text(raster(r_max), digits = 3)

```

:::
